### 2015-01-08 prelim/oral_prelim.tex
updates to prelim

\documentclass{beamer}
% \usepackage{beamerthemesplit}
\usetheme{Madrid}
% \usepackage{pstricks}
\usepackage{graphicx}
\usepackage{mdwlist}
\usepackage{lineno, hyperref}
\usepackage{stackrel}

\usepackage{amssymb,latexsym,amsmath,amsthm,bbm}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage{multirow}
\usepackage{verbatim}
\usepackage{alltt}
\usepackage{mycommands}

@@

\begin{frame}{Motivation}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Average behavior is important to understand, but it does not paint the whole picture
    \begin{itemize}
      \item e.g. When constructing river levees, engineers need to be able to estimate a 100-year or 1000-year flood levels
      \item e.g. Probability of exceeding a certain threshold level
    \end{itemize}
    \item Estimating the probability of rare events is challenging because these events are, by definition, rare
    \item Spatial extremes is promising because it borrows informaation across space
    \item Spatial extremes is also useful for estimating probability of extremes at sites without data
  \end{itemize}
\end{frame}

@@

\begin{frame}{Motivation}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Ozone compliance for Clean Air Act (EPA)
    \begin{itemize}
      \item Annual fourth-highest daily maximum 8-hour concentration, averaged over 3 years, not to exceed 75 ppb
      \item Annual fourth-highest is the 99th percentile for the year
      \item Common objectives are
      \begin{itemize}
        \item To interpolate to unmonitored sites
        \item Detect changes in extremes over time
        \item Study factors that lead to extreme events
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

@@

  \column{.55\linewidth}
  \begin{figure}
    \includegraphics[width=1\linewidth, trim=0 0.5in 0 1in]{./plots/define_extreme.pdf}
    \caption{Hypothetical monthly maximums recorded over a three years}
    \end{figure}
\end{columns}
\end{frame}

\begin{frame}{Standard non-spatial analysis: Block maxima}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Let $X_1, \ldots, X_n$ be i.i.d.
    \item Consider the block maximum $M_n = \max(X_1, \ldots, X_n)$
    \item If there exist normalizing sequences $a_n > 0$ and $b_n \in \calR$ such that
    \begin{align*}
      \frac{ M_n - b_n }{ a_n } \converged G(z)
    \end{align*}
    then $G(z)$ follows a generalized extreme value distribution (GEV) (Falk et al., 2011)
    \item This motivates the use of the GEV for block max data
  \end{itemize}
\end{frame}

\begin{frame}{Standard non-spatial analysis: Block maxima}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item GEV distribution
    \begin{align*}
      G(y) = \Pr(Y < y) = \left\{  \begin{array}{ll}
        \exp\left\{ -\left[ 1 + \xi \left( \frac{ y - \mu }{ \sigma } \right) \right]^{ -1 / \xi} \right\} & \quad \xi \neq 0 \\[0.5em]
        \exp \left\{ -\exp \left( - \frac{ y - \mu }{ \sigma} \right) \right\} & \quad \xi = 0
      \end{array}\right.
    \end{align*}
    where
    \begin{itemize}
      \item $\mu \in \calR$ is a location parameter
      \item $\sigma > 0$ is a scale parameter
      \item $\xi \in \calR$ is a shape parameter
      \begin{itemize}
        \item Unbounded above if $\xi \ge 0$
        \item Bounded above by $(\mu - \sigma) / \xi$ when $\xi < 0$
      \end{itemize}
    \end{itemize}
    \item Challenges:
    \begin{itemize}
      \item Lose information by only considering maximum in a block
      \item Underlying data may not be i.i.d.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Standard non-spatial analysis: Peaks over threshold}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Let $X_1, \ldots, X_n \sim F$
    \item If there exist normalizing sequences $a_t > 0$ and $b_t \in \calR$ such that for any $x \ge 0$, as $T \rightarrow \infty$
    \begin{align*}
      \Pr\left(\frac{X - b_t}{a_t} > t \mid X > T\right) \converged H(x),
    \end{align*}
    where $T$ is a thresholding value, then $H(x)$ follows a generalized Pareto distribution (GPD) (Balkema and de Haan, 1974)
  \end{itemize}
\end{frame}

\begin{frame}{Standard non-spatial analysis: Peaks over threshold}
  \begin{itemize}  \setlength{\itemsep}{1em}
    \item Select a threshold, $T$, and use the Generalized Pareto distribution to model the exceedances
    \begin{align*}
      H(y) = P(Y < y) = \left\{ \begin{array}{ll}
        1 - \left[1 - \xi \left( \frac{ y - T }{ \sigma } \right) \right]^{-1 / \xi} & \quad \xi \neq 0 \\[0.5em]
        1 - \exp \left\{ \frac{ y - T }{ \sigma} \right\} & \quad \xi = 0
      \end{array}\right.
    \end{align*}
    where
    \begin{itemize}
      \item $\sigma > 0$ is a scale parameter
      \item $\xi \in \calR$ is a shape parameter
      \begin{itemize}
        \item Unbounded above if $\xi \ge 0$
        \item Bounded above by $(T - \sigma) / \xi$ when $\xi < 0$
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Standard non-spatial analysis: Peaks over threshold}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item The GPD is related to GEV distribution through
    \begin{align*}
      H(y) = 1 + \log[G(y)], \quad y \ge T
    \end{align*}
    \item Challenges:
    \begin{itemize}
      \item Sensitive to threshold selection
      \item Temporal dependence between observations (e.g. flood levels don't dissipate overnight)
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Max-stable processes for spatial data}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Consider i.i.d. spatial processes $x_j(\bs)$, $j = 1, \ldots, J$
    \item Let $M_J(\bs) = \bigvee_{j=1}^J x_j(\bs_i)$ be the block max at site $\bs$
    \item If there exists normalizing sequences $a_J(\bs)$ and $b_J(\bs)$ such that for all sites, $\bs_i, i = 1, \ldots, d$,
    \begin{align*}
      a_J^{-1}(\bs) \left\{ M_J(\bs) - b_J(\bs) \right\} \converged Y(\bs)
    \end{align*}
    which has a non-degenerate distribution, then $Y(\bs)$ is a max-stable process
    \item Therefore, max-stable processes are the standard for block maxima
  \end{itemize}
\end{frame}

\begin{frame}{Multivariate representations}
  \begin{itemize}
    \item Marginally at each site, observations follow a GEV distribution
    \item For a finite collection of sites the multivariate representation for the GEV (mGEV) is
    \begin{align*}
      \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp(-V(\bz))\\
            V(\bz)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
    \end{align*}
    where
    \begin{itemize}
      \item $\Delta_d = \{ \bw \in \calR^d_+ \mid w_1 + \cdots + w_d = 1\}$
      \item $H$ is a probability measure on $\Delta_d$
      \item $\int_{\Delta_d}w_i H(\ddd w) = 1 / d$ for $i = 1, \ldots, d$.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Multivariate GEV challenges}
  \begin{itemize}
    \item Only a few closed-form expressions for $V(\bz)$ exist (Stephenson, 2003)
    \item Two common forms for $V(\bz)$:
    \begin{itemize}
      \item Symmetric logistic
      \begin{align*}
        V(\bz) = \left[\sum_{i = 1}^n \left( \frac{ 1 }{ z_i } \right)^{1/\alpha}\right]^\alpha
      \end{align*}
      \item Asymmetric logistic
      \begin{align*}
        V(\bz) = \sum_{l = 1}^L \left[\sum_{i = 1}^n \left(\frac{w_{il}}{z_i} \right)^{1 / \alpha_l} \right]^{\alpha_l}
      \end{align*}
      where $w_{il} \in [0, 1]$ and $\sum_{l = 1}^L w_{il} = 1$.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Multivariate peaks over threshold}
  \begin{itemize}
    \item Few existing methods
    \item Often use max-stable methods due to the relationship between GEV and GPD
    \item Joint distribution function given by Falk et al. (2011)
    \begin{align*}
      H(\bz) = 1 - V(\bz)
    \end{align*}
    where $V(\bz)$ is defined as in the GEV
  \end{itemize}
\end{frame}

@@

\begin{frame}{Existing challenges}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Multivariate max-stable and GPD models have nice features, but they are
    \begin{itemize}
      \item Computationally challenging (e.g,  the asymmetric logistic has $2^{n-1}(n + 2) - (2n + 1)$ free parameters)
      \item Joint distribution only available in low dimensions
    \end{itemize}
    \item Some recent approaches
    \begin{itemize}
      \item Bayesian hierarchical model (Reich and Shaby, 2012)
      \item Pairwise likelihood approach (Huser and Davison, 2014)
    \end{itemize}
    \item Many opportunities to explore new methods
  \end{itemize}
\end{frame}

\begin{frame}{Three principal contributions}
  \begin{enumerate}[1] \setlength{\itemsep}{1em}
    \item A spatio-temporal model with flexible tails, asymptotic spatial dependence, and computation on the order of Gaussian models for large space-time datasets
    \item Predicting rare binary events with a spatially dependent generalized extreme-value link function
    \item A Bayesian hierarchical model to allow for non-stationary covariance in extreme value models
  \end{enumerate}
\end{frame}

\begin{frame}{Spatiotemporal modeling for extreme values}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Model objectives:
    \begin{itemize}
      \item Marginal distribution with a flexible tail
      \begin{itemize}
        \item Allow for asymmetric distributions
        \item Allow for heavy tails
      \end{itemize}
      \item Asymptotic spatial dependence
      \item Computation on the order of Gaussian models for large space-time datasets
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Gaussian spatial model}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item In geostatistics $Y(\bs)$ are often modeled using a Gaussian process with mean function $\mu(\bs)$ and covariance function $\rho(h)$.
    \item Model properties
    \begin{itemize}
      \item Nice computing properties (closed-form likelihood)
      \item For a Gaussian spatial model $\lim_{c \rightarrow \infty} \chi(h) = 0$ regardless of the strength of the correlation in the bulk of the distribution
      \item Tail is not flexible
      \begin{itemize}
        \item Light-tailed
        \item Symmetric
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Spatial skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item A more flexible alternative is the spatial skew-$t$ process (Zhang and El-Shaarawi, 2012)
    \begin{align*}
      Y(\bs) = \bX(\bs)\bbeta + \lambda |z| + v(\bs)
    \end{align*}
    where
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item $\lambda \in \calR$ controls the skewness
      \item $z \sim N(0, \sigma^2)$ is a random effect
      \item $v(\bs)$ is a Gaussian process with variance $\sigma^2$ and \Matern correlation
      \item $\sigma^2 \sim \text{IG}(a, b)$
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Spatial skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item \alert{Conditioned} on $z$ and $\sigma^2$, $Y(\bs)$ is a Gaussian spatial model
    \item Can use standard geostatistical methods to fit this model
    \item Predictions can be made through Kriging
  \end{itemize}
\end{frame}

\begin{frame}{Spatial skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item \alert{Marginalizing} over $z$ and $\sigma^2$ (via MCMC),
    \begin{align*}
      Y(\bs) \sim \text{skew-t}(\bX(\bs), \bOmega, \alpha, \text{df}=2a)
    \end{align*}
    where
    \begin{itemize}
      \item $\bX(\bs) \bbeta$ is the location
      \item $\bOmega = \frac{1}{ab} \bar{\bOmega}$ is a correlation matrix
      \item $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T)$
      \item $\bSigma$ is a postive definite correlation matrix
      \item $\alpha = \lambda( 1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ controls the skewness
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Spatial skew-$t$ distribution properties}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Model properties:
    \begin{itemize}
      \item Flexible tail
      \begin{itemize}
        \item Skewness controlled by $\lambda$
        \item Weight of tails controlled by degrees of freedom $2a$
      \end{itemize}
      \item For a skew-$t$ distribution $\lim_{c \rightarrow \infty} \chi(h) > 0$ (Padoan, 2011)
      \item Computation that is on the order of Gaussian computation
    \end{itemize}
    \item Challenge: $\chi(h) > 0$ as $h \rightarrow \infty$ (Padoan, 2011)
    \begin{itemize}
      \item This occurs because all observations (near and far) share the same $z$ and $\sigma^2$
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Extension of the skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Skew-$t$ distribution addresses two modeling concerns
    \begin{itemize}
      \item Extremal dependence
      \item Reasonable computing
    \end{itemize}
    \item Our contribution is to extend the skew-$t$
    \begin{itemize}
      \item Censoring to focus on extreme observations
      \item Partitioning to address long-range dependence
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Random partition}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Within each partition $Y(\bs)$ has the same MV skew-t distribution as before
    \item Across partitions $Y(\bs)$ are asymptotically independent, but still correlated through $v(\bs)$
    \item New expression for $\chi(h)$
    \begin{align*}
      \chi(h) = \pi(h) \chi_{\text{skew-}t}(h)
    \end{align*}
    where
    \begin{itemize}
      \item $\pi(h)$ is the probability two sites are in the same partition
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Random partition skew-$t$ model}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item This new model is called a random partition skew-$t$ model, and it has the properties we desire
    \begin{itemize}
      \item Marginal distribution with flexible tails
      \begin{itemize}
        \item $\lambda$ term allows for asymmetry
        \item Degrees of freedom control heavy vs light tails
      \end{itemize}
      \item Asymptotic spatial dependence for that decays with distance between sites through partitioning
      \item Computation is on the order of Gaussian models for large space-time datasets
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{MCMC details}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Three main steps
    \begin{enumerate}[1.]
      \item Impute censored data below $T$
      \item Update parameters with standard random walk Metropolis Hastings or Gibbs sampling
      \item Make spatial predictions
    \end{enumerate}
    \item Priors are selected to be conjugate when possible
  \end{itemize}
\end{frame}

>>>>

\documentclass{beamer}
% \usepackage{beamerthemesplit}
\usetheme{Madrid}
% \usepackage{pstricks}
\usepackage{graphicx}
\usepackage{mdwlist}
\usepackage{lineno, hyperref}
\usepackage{stackrel}
\usepackage{changepage}

\usepackage{amssymb,latexsym,amsmath,amsthm,bbm}
\usepackage{hyperref}
\usepackage{tikz}
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage[authoryear]{natbib}
\usepackage{multirow}
\usepackage{verbatim}
\usepackage{alltt}
\usepackage{mycommands}

@@

\begin{frame}{Motivation}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Average behavior is important to understand, but it does not paint the whole picture
    \begin{itemize}
      \item e.g. When constructing river levees, engineers need to be able to estimate a 100-year or 1000-year flood levels
      \item e.g. Probability of ambient air pollution exceeding a certain threshold level
    \end{itemize}
    \item Estimating the probability of rare events is challenging because these events are, by definition, rare
    \item Spatial extremes is promising because it borrows informaation across space
    \item Spatial extremes is also useful for estimating probability of extremes at sites without data
  \end{itemize}
\end{frame}

@@

\begin{frame}{Motivation}
\begin{adjustwidth}{1em}{0em}
  Ozone compliance for Clean Air Act (EPA) \vspace{1em}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Annual fourth-highest daily maximum 8-hour concentration, averaged over 3 years, not to exceed 75 ppb
    \item Annual fourth-highest is the 99th percentile for the year
    \item Common objectives are \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item To interpolate to unmonitored sites
      \item Detect changes in extremes over time
      \item Study meterological conditions that lead to extreme events
    \end{itemize}
  \end{itemize}
\end{adjustwidth}
\end{frame}

@@

  \column{.55\linewidth}
  \begin{figure}
    \includegraphics[width=1\linewidth, trim=0 0.5in 0 1in]{./plots/define_extreme.pdf}
    \caption{Hypothetical monthly data}
    \end{figure}
\end{columns}
\end{frame}

\begin{frame}{Non-spatial analysis: Block maxima}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Let $X_1, \ldots, X_n$ be i.i.d.
    \item Consider the block maximum $M_n = \max(X_1, \ldots, X_n)$
    \item If there exist normalizing sequences $a_n > 0$ and $b_n \in \calR$ such that
    \begin{align*}
      \frac{ M_n - b_n }{ a_n } \converged G(z)
    \end{align*}
    then $G(z)$ follows a generalized extreme value distribution (GEV) (Falk et al., 2011)
    \item This motivates the use of the GEV for block maximum data
  \end{itemize}
\end{frame}

\begin{frame}{Non-spatial analysis: Block maxima}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item GEV distribution
    \begin{align*}
      G(y) = \Pr(Y < y) = \left\{  \begin{array}{ll}
        \exp\left\{ -\left[ 1 + \xi \left( \frac{ y - \mu }{ \sigma } \right) \right]^{ -1 / \xi} \right\} & \quad \xi \neq 0 \\[0.5em]
        \exp \left\{ -\exp \left( - \frac{ y - \mu }{ \sigma} \right) \right\} & \quad \xi = 0
      \end{array}\right.
    \end{align*}
    where
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item $\mu \in \calR$ is a location parameter
      \item $\sigma > 0$ is a scale parameter
      \item $\xi \in \calR$ is a shape parameter
      \begin{itemize}
        \item Unbounded above if $\xi \ge 0$
        \item Bounded above by $(\mu - \sigma) / \xi$ when $\xi < 0$
      \end{itemize}
    \end{itemize}
    \item Challenges:
    \begin{itemize}
      \item Lose information by only considering maximum in a block
      \item Underlying data may not be i.i.d.
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Non-spatial analysis: Peaks over threshold}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Let $X_1, \ldots, X_n \sim F$
    \item If there exist normalizing sequences $a_t > 0$ and $b_t \in \calR$ such that for any $x \ge 0$, as $T \rightarrow \infty$
    \begin{align*}
      \Pr\left(\frac{X - b_t}{a_t} > x \mid X > T\right) \converged H(x),
    \end{align*}
    where $T$ is a thresholding value, then $H(x)$ follows a generalized Pareto distribution (GPD) (Balkema and de Haan, 1974)
  \end{itemize}
\end{frame}

\begin{frame}{Non-spatial analysis: Peaks over threshold}
\begin{adjustwidth}{1em}{0em}
  Select a threshold, $T$, and use the GPD to model the exceedances
  \begin{align*}
    H(y) = P(Y < y) = \left\{ \begin{array}{ll}
      1 - \left[1 - \xi \left( \frac{ y - T }{ \sigma } \right) \right]^{-1 / \xi} & \quad \xi \neq 0 \\[0.5em]
      1 - \exp \left\{ \frac{ y - T }{ \sigma} \right\} & \quad \xi = 0
    \end{array}\right.
  \end{align*}
  where
  \begin{itemize} \setlength{\itemsep}{0.25em}
    \item $\sigma > 0$ is a scale parameter
    \item $\xi \in \calR$ is a shape parameter
    \begin{itemize}
      \item Unbounded above if $\xi \ge 0$
      \item Bounded above by $(T - \sigma) / \xi$ when $\xi < 0$
    \end{itemize}
  \end{itemize}
\end{adjustwidth}
\end{frame}

\begin{frame}{Non-spatial analysis: Peaks over threshold}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item The GPD is related to GEV distribution through
    \begin{align*}
      H(y) = 1 + \log[G(y)], \quad y \ge T
    \end{align*}
    \item Challenges: \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Sensitive to threshold selection
      \item Temporal dependence between observations (e.g. flood levels don't dissipate overnight)
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Max-stable processes for spatial data}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Consider i.i.d. spatial processes $x_j(\bs)$, $j = 1, \ldots, J$
    \item Let $M_J(\bs) = \bigvee_{j=1}^J x_j(\bs_i)$ be the block maximum at site $\bs$
    \item If there exists normalizing sequences $a_J(\bs)$ and $b_J(\bs)$ such that for all sites, $\bs_i, i = 1, \ldots, d$,
    \begin{align*}
      a_J^{-1}(\bs) \left\{ M_J(\bs) - b_J(\bs) \right\} \converged Y(\bs)
    \end{align*}
    which has a non-degenerate distribution, then $Y(\bs)$ is a max-stable process
    \item Therefore, max-stable processes are the standard model for block maxima
  \end{itemize}
\end{frame}

\begin{frame}{Multivariate representations}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Marginally at each site, observations follow a GEV distribution
    \item For a finite collection of sites the multivariate representation for the GEV (mGEV) is
    \begin{align*}
      \Pr(\bZ \le \bz)  &= G^*(\bz) = \exp(-V(\bz))\\
            V(\bz)    &= d \int_{\Delta_d} \bigvee_{i = 1}^d \frac{w_i}{z_i} H(\ddd w)
    \end{align*}
    where
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item $\Delta_d = \{ \bw \in \calR^d_+ \mid w_1 + \cdots + w_d = 1\}$
      \item $H$ is a probability measure on $\Delta_d$
      \item $\int_{\Delta_d}w_i H(\ddd w) = 1 / d$ for $i = 1, \ldots, d$
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Multivariate GEV challenges}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Only a few closed-form expressions for $V(\bz)$ exist (Stephenson, 2003)
    \item Two common forms for $V(\bz)$:
    \begin{itemize}
      \item Symmetric logistic
      \begin{align*}
        V(\bz) = \left[\sum_{i = 1}^n \left( \frac{ 1 }{ z_i } \right)^{1/\alpha}\right]^\alpha
      \end{align*}
      \item Asymmetric logistic
      \begin{align*}
        V(\bz) = \sum_{l = 1}^L \left[\sum_{i = 1}^n \left(\frac{w_{il}}{z_i} \right)^{1 / \alpha_l} \right]^{\alpha_l}
      \end{align*}
      where $w_{il} \in [0, 1]$ and $\sum_{l = 1}^L w_{il} = 1$
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Multivariate peaks over threshold}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Few existing methods
    \item Often use max-stable methods due to the relationship between GEV and GPD
    \item Joint distribution function given by Falk et al. (2011)
    \begin{align*}
      H(\bz) = 1 - V(\bz)
    \end{align*}
    where $V(\bz)$ is defined as in the GEV
  \end{itemize}
\end{frame}

@@

\begin{frame}{Existing challenges}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Multivariate max-stable and GPD models have nice features, but they are
    \begin{itemize}
      \item Computationally challenging (e.g,  the asymmetric logistic has $2^{n-1}(n + 2) - (2n + 1)$ free parameters)
      \item Joint density only available in low dimensions
    \end{itemize}
    \item Some recent approaches
    \begin{itemize}
      \item Bayesian hierarchical model (Reich and Shaby, 2012)
      \item Pairwise likelihood approach (Huser and Davison, 2014)
    \end{itemize}
    \item Many opportunities to explore new methods
  \end{itemize}
\end{frame}

\begin{frame}{Three principal contributions}
\begin{adjustwidth}{1em}{0em}
  \begin{enumerate}[1.] \setlength{\itemsep}{1.5em}
    \item A spatio-temporal model with flexible tails, asymptotic spatial dependence, and computation on the order of Gaussian models for large space-time datasets
    \item Predicting rare binary events with a spatially dependent generalized extreme-value link function
    \item A Bayesian hierarchical model to allow for non-stationary covariance in extreme value models
  \end{enumerate}
\end{adjustwidth}
\end{frame}

\begin{frame}{Spatiotemporal modeling for extreme values}
\begin{adjustwidth}{1em}{0em}
  Model objectives: \vspace{1em}
  \begin{itemize} \setlength{\itemsep}{1em} \setlength{\itemindent}{1em}
    \item Marginal distribution at each site with a flexible tail
    \begin{itemize}
      \item Allow for asymmetric distributions
      \item Allow for heavy tails
    \end{itemize}
    \item Asymptotic spatial dependence
    \item Computation on the order of Gaussian models for large space-time datasets
  \end{itemize}
\end{adjustwidth}
\end{frame}

\begin{frame}{Gaussian spatial model}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item In geostatistics, $Y(\bs)$ are often modeled using a Gaussian process with mean function $\mu(\bs)$ and covariance function $\rho(h)$.
    \item Model properties
    \begin{itemize}
      \item Nice computing properties (closed-form likelihood)
      \item For a Gaussian spatial model $\chi(h) = 0$ regardless of the strength of the correlation in the bulk of the distribution
      \item Tail is not flexible
      \begin{itemize}
        \item Light-tailed
        \item Symmetric
      \end{itemize}
    \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Spatial skew-$t$ distribution}
\begin{adjustwidth}{1em}{0em}
  A more flexible alternative is the spatial skew-$t$ process (Zhang and El-Shaarawi, 2012)
  \begin{align*}
    Y(\bs) = \bX(\bs)\bbeta + \lambda |z| + v(\bs)
  \end{align*}
  where
  \begin{itemize} \setlength{\itemsep}{0.25em}
    \item $\lambda \in \calR$ controls the skewness
    \item $z \sim N(0, \sigma^2)$ is a random effect
    \item $v(\bs)$ is a Gaussian process with variance $\sigma^2$ and \Matern correlation
    \item $\sigma^2 \sim \text{IG}(a, b)$
  \end{itemize}
\end{adjustwidth}
\end{frame}

\begin{frame}{Spatial skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item \alert{Conditioned} on $z$ and $\sigma^2$, $Y(\bs)$ is a Gaussian spatial model
    \item Standard geostatistical methods apply
    \item Predictions can be made through Kriging
  \end{itemize}
\end{frame}

\begin{frame}{Spatial skew-$t$ distribution}
\begin{adjustwidth}{1em}{0em}
  \alert{Marginalizing} over $z$ and $\sigma^2$ (via MCMC),
  \begin{align*}
    Y(\bs) \sim \text{skew-t}(\bX(\bs), \bOmega, \alpha, \text{df}=2a)
  \end{align*}
  where
  \begin{itemize} \setlength{\itemsep}{0.25em}
    \item $\bX(\bs) \bbeta$ is the location
    \item $\bOmega = \frac{1}{ab} \bar{\bOmega}$ is a correlation matrix
    \item $\bar{\bOmega} = (\bSigma + \lambda^2 \bOne \bOne^T)$
    \item $\bSigma$ is a postive definite correlation matrix
    \item $\alpha = \lambda( 1 + \lambda^2 \bOne^T \bSigma^{-1} \bOne)^{-1/2} \bOne^T \bSigma^{-1}$ controls the skewness
  \end{itemize}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{Spatial skew-$t$ distribution properties}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Model properties: \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Flexible tail
      \begin{itemize}
        \item Skewness controlled by $\lambda$
        \item Weight of tails controlled by degrees of freedom $2a$
      \end{itemize}
      \item For a skew-$t$ distribution $\lim_{c \rightarrow \infty} \chi(h) > 0$ (Padoan, 2011)
      \item Computation that is on the order of Gaussian computation
    \end{itemize}
    \item Challenge: Long-range dependence occurs because all observations (near and far) share the same $z$ and $\sigma^2$ (Padoan, 2011)
    \begin{align*}
      \lim_{h \rightarrow \infty} \chi(h) > 0
    \end{align*}
  \end{itemize}
\end{frame}

\begin{frame}{Extension of the skew-$t$ distribution}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Skew-$t$ distribution addresses two modeling concerns \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Extremal dependence
      \item Reasonable computing
    \end{itemize}
    \item Our contribution is to extend the skew-$t$ \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item Censoring to focus on extreme observations
      \item Partitioning to address long-range dependence
    \end{itemize}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Random partition}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Within each partition, $Y(\bs)$ has the same MV skew-t distribution as before
    \item Across partitions $Y(\bs)$ are asymptotically independent, but still correlated through $v(\bs)$
    \item New expression for $\chi(h)$
    \begin{align*}
      \chi(h) = \pi(h) \chi_{\text{skew-}t}(h)
    \end{align*}
    where $\pi(h)$ is the probability two sites are in the same partition
  \end{itemize}
\end{frame}

\begin{frame}{Proof that $\lim_{h \rightarrow \infty} \chi(h) = 0$}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Let $A$ be the area between two sites $\bs$ and $\bt$
    \item Let $N(A)$ be the number of knots in $A$
    \item Assume that $N(A) \sim \text{HPP}[\mu(A)]$, where \vspace{0.25em}
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item HPP$[\mu]$ is a homogeneous poisson process with intensity $\mu$
      \item $\mu(\cdot)$ is an intensity measure defined on $A$
      \item $\lim_{h \rightarrow \infty} \mu(A) = \infty$
    \end{itemize}
    \begin{align*}
      \Pr[N(A) = k] = \frac{ \mu(A)^k \exp\{-\mu(A)\}}{k!}
    \end{align*}
  \end{itemize}
\end{frame}

\begin{frame}{Proof that $\lim_{h \rightarrow \infty} \chi(h) = 0$}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item For finite $k$, $\lim_{h \rightarrow \infty} P[N(A) = k] = 0$
    \item As $N(A)$ increases, $\pi(h)$ decreases because partition is defined by closest knots, so
    \begin{align*}
      \lim_{h \rightarrow \infty} \chi(h) = \lim_{h \rightarrow \infty} \pi(h) \chi_{\text{skew-}t} (h) = 0
    \end{align*}
  \end{itemize}
\end{frame}

@@

\begin{frame}{Random partition skew-$t$ model}
\begin{adjustwidth}{1em}{0em}
  This new model is called a random partition skew-$t$ model, and it has all the properties we desire \vspace{1em}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Marginal distribution with flexible tails \vspace{0.5em}
    \begin{itemize} \setlength{\itemsep}{0.5em}
      \item $\lambda$ term allows for asymmetry
      \item Degrees of freedom control heavy vs light tails
    \end{itemize}
    \item Asymptotic spatial dependence for that decays with distance between sites through partitioning
    \item Computation is on the order of Gaussian models for large space-time datasets
  \end{itemize}
\end{adjustwidth}
\end{frame}

@@

\begin{frame}{MCMC details}
  \begin{itemize} \setlength{\itemsep}{1em}
    \item Three main steps \vspace{0.5em}
    \begin{enumerate}[1.] \setlength{\itemsep}{0.5em}
      \item Impute censored data below $T$
      \item Update parameters with standard random walk Metropolis Hastings or Gibbs sampling
      \item Make spatial predictions
    \end{enumerate}
    \item Priors are selected to be conjugate when possible
  \end{itemize}
\end{frame}

\begin{frame}{Simulation study}
\begin{adjustwidth}{1em}{0em}
  6 different data settings \vspace{0.5em}
  \begin{enumerate}[1.] \setlength{\itemsep}{0.5em}
    \item Gaussian, $K = 1$ partition
    \item Symmetric-$t$, $K = 1$ partition
    \item Symmetric-$t$, $K = 5$ partitions
    \item Skew-$t$, $K = 1$ partition
    \item Skew-$t$, $K = 5$ partitions
    \item Max-stable \vspace{0.25em}
    \begin{itemize} \setlength{\itemsep}{0.25em}
      \item Marginally: GEV$(\mu=1, \sigma=1, \xi=0.2)$
      \item Dependence function: asymmetric logistic with $\alpha = 0.5$
    \end{itemize}
  \end{enumerate}
\end{adjustwidth}
\end{frame}
